Ensuring Fairness with Transparent Auditing of Quantitative Bias in AI Systems

### 2024-08-29 slides/2024_8_30.tex
Added more description
Added slides

\begin{frame}
    \frametitle{Predicate}
    We first introduce the idea of a predicate.
    \begin{itemize}
        \item A \emph{predicate} is a decider that determines
        if some condition holds for some given input.
        \item Formally, it is a function that takes
        some type $\alpha$ and returns True or False.
        \item For example, an ``$\text{adult}(x)$'' predicate might
        take a person $x$'s data and returns if they are over 18
        year old.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Abstraction}
    Our framework tool abstracts the fairness measures:
    \begin{itemize}
        \item A \emph{row} is a lookup table or dictionary.
        $r_n(\text{``sex''}) = \text{``Female''}$ means $r_n$'s sex is female.
        \item A \emph{privileged predicate} $\priv$ takes a row and determines if
        it belongs to the privileged group.
        For example, $\priv(r_i) := r_i(\text{``race''}) == \text{``Caucasian''}$
        means the privileged group is those with race being Caucasian.
        \item A \emph{positive predicate} $\pos$ takes a row and determines if its
        prediction is positive. For example,
        $\pos(r_i) := \mathtt{int}(r_i(\text{``score''})) > 7$ means a
        row's prediction is positive if its score is greater than 7.
        \item A \emph{ground truth predicate} $\tru$ takes a row and gives the ground
        truth of the result. For example,
        $\tru(r_i) := r_i(\text{``recid''}) == \text{True}$ indicates a
        row's actual recidivism.
    \end{itemize}
\end{frame}

@@

    This slides is available at \url{https://github.com/RexYuan/Shu}.
\end{frame}

>>>>

\begin{frame}
    \frametitle{Predicate}
    We first introduce the idea of a predicate.
    \begin{itemize}
        \item A \emph{predicate} is a decider that determines
        if some condition holds for some given input.
        \item For example, an ``$\text{adult}(x)$'' predicate might
        take a person $x$'s data and returns if they are over 18
        year old.
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Predicate}
    \[
    \begin{array}{rcl}
    \text{adult}(r_i) & := & r_i(\text{``Age''}) > 18 \\
    \text{rich}(r_i) & := & r_i(\text{``Income''}) > 3000
    \end{array}
    \]
    \begin{table}
        \begin{tabular}{|c|c|c|c|c|}
            \hline
                  & Age & Income & $\text{adult}(r_i)$ & $\text{rich}(x)$ \\
            \hline
            $r_1$ & 14 & \$10000 & False & True  \\
            $r_2$ & 17 & \$1000 & False & False  \\
            $r_3$ & 45 & \$1500 & True & False  \\
            $r_4$ & 40 & \$2000 & True & False  \\
            $r_5$ & 32 & \$5000 & True & True  \\
            \hline
        \end{tabular}
        \caption{Example Dataset with Predicates}
    \end{table}
\end{frame}

\begin{frame}
    \frametitle{Abstraction}
    Our framework tool abstracts the fairness measures:
    \begin{itemize}
        \item A \emph{row} is a lookup table or dictionary.
        $r_n(\text{``sex''}) = \text{``Female''}$ means $r_n$'s sex is female.
        \item A \emph{privileged predicate} $\priv$ takes a row and determines if
        it belongs to the privileged group.
        For example, $\priv(r_i) := r_i(\text{``race''}) == \text{``Caucasian''}$
        means the privileged group is those with race being Caucasian.
        \item A \emph{positive predicate} $\pos$ takes a row and determines if its
        prediction is positive. For example,
        $\pos(r_i) := r_i(\text{``score''}) > 7$ means a
        row's prediction is positive if its score is greater than 7.
        \item A \emph{ground truth predicate} $\tru$ takes a row and gives the ground
        truth of the result. For example,
        $\tru(r_i) := r_i(\text{``recid''}) == \text{True}$ indicates a
        row's actual recidivism.
    \end{itemize}
\end{frame}

@@

    Slides: \url{https://github.com/RexYuan/Shu}.

    Package: \url{https://pypi.org/project/fairness-checker}.
\end{frame}
